NDNS multilevel LCA CW2CB2 or CW2CB3

### 2018-08-19 paper/Chapters/Chapter4.tex
updated discussion done

Firstly, we ignored the order of observation days in the MLCA models. The food consumption diaries were accomplished by participants for at least 3 out of 4 consecutive days. In the multilevel analyses, these 3 or 4 days' observations were treated as if they were exchangeable in the models. Since one's diet might change according to season, day of the week (weekdays or weekends), or sometimes depends on what they had consumed the day before, exchangeability of their diaries is a strong assumption and cannot be overcome by the MLCA models adopted here. Other statistical techniques, which could take the order or the longitudinal nature of the data into account, such as repeated measures latent class analysis (RMLCA), latent transition analysis (LTA) \parencite{collins2010latent}, or latent class growth analysis (LCGA) \parencite{davidian2008growth,jung2008introduction,andruff2009latent} are not applicable for the NDNS RP dataset. For the purpose of maintaining the response rate, flexibility was allowed for the NDNS RP participants in their choice of which day to begin their diary.  Additionally, these models require that the 3 or 4 repeated measurements of food diaries should be recorded at the same time points longitudinally. LCGA models (also called growth mixture modelling) will need to model the change of the odds of the probability over time as a function (quadratic, or cubic) which is apparently not the objective in the current project.

Secondly, the classification of individuals to latent carbohydrate eating classes were defined by maximum posterior probability assignment rule. This approach assigns individuals to the class for which they have the highest posterior probability of membership \parencite{nagin2005group}. The other approach is multiple pseudo-class draws \parencite{wang2005residual}, which was proposed in attempt to account for uncertainty in class assignment. However, the maximum probability rule is still believed to be able to minimize the number of incorrect assignments \parencite{goodman20071}. A Monte Carlo simulation study \parencite{bray2015eliminating} also demonstrated that maximum probability assignment is less biased than multiple pseudo-class draws. Moreover, this simulation study also found that an inclusive LCA (i.e. LCA with covariates) would perform better, and reduce bias. Since whether this advantage can be extended to MLCA in the current project is unknown, and also due to the reason that the associations between the latent classes and distal outcomes (hypertension, and BMI) in our analyses were mostly still under exploration, therefore, the model for LCA with covariates can never be well-known before our analyses. Thus, as a first step, also given the complexity of the NDNS RP dataset itself, we chose to conduct the MLCA modellings without any covariate in either day level or individual level models. Future studies may be advised to consider incorporating other predictors in the MLCA models to see whether the classifications in both levels can be improved or not.

Thirdly, information regarding the detailed occupations of the participants was not available, so we could not exclude those who had a shift work involving night work. But considering that completing a 4-consecutive-day food diary would be a large burden for the participants, those had a shift work involving night work might possibly performed relatively lower compliance and could not complete their diary for at least 3 days. But this cannot be verified in the NDNS RP dataset. If people working with shift work were included in the NDNS RP sample, the classification in both levels as well as their associations with either hypertension or BMI would be biased. Furthermore, under-reporting is also a concern in the diaries collected. To evaluate the influence of under-reporting in the NDNS RP sample, study team of NDNS RP conducted doubly labelled water (DLW) sub-study within the survey. The details of this sub-study can be accessed from the Appendix X of the Official Reports provided by Public Health England \parencite{bates2014national,roberts2018national,NDNSofficial}.

\begin{itemize}
	\item dietary report under estimation
	\item findings may not be generalizable to populations from other countries.
\end{itemize}

@@

We have successfully defined carbohydrate eating patterns in the general population in the UK adults using the NDNS RP database in both observation day level and participant level. Low carbohydrate eaters turned to have more energy that contributed by both fat and alcohol. Moderate carbohydrate eaters consumed the lowest total energy, while they had the tendency of having meals later in time-of-day. High carbohydrate eaters consumed most of their carbohydrate as well as energy earlier in time-of-day. These dietary patterns specifically for carbohydrate intake were found to be differed by timing, quantity and resources of energy consumption. Compared with low carbohydrate eaters, men had moderate carbohydrate eating pattern may associated with a lower prevalence of hypertension, women in this latent class who lived alone may associated with a larger waist circumferences. Among women who lived with partners, high carbohydrate eating pattern was associated with both lower BMI and smaller waist circumferences. Future studies exploring how such carbohydrate eating patterns may relate longitudinally to change of obesity, hypertension and diabetes incidence and other health outcomes are needed.

>>>>

Firstly, we ignored the order of observation days in the MLCA models. The food consumption diaries were accomplished by participants for at least 3 out of 4 consecutive days. In the multilevel analyses, these 3 or 4 days' observations were treated as if they were exchangeable in the models. Since one's diet might change according to season, day of the week (weekdays or weekends), or sometimes depends on what they had consumed the day before, exchangeability of their diaries is a strong assumption and cannot be overcome by the MLCA models adopted here. Other statistical techniques, which could take the order or the longitudinal nature of the data into account, such as repeated measures latent class analysis (RMLCA), latent transition analysis (LTA) \parencite{collins2010latent}, or latent class growth analysis (LCGA) \parencite{davidian2008growth,jung2008introduction,andruff2009latent} are not applicable for the NDNS RP dataset. For the purpose of maintaining the response rate, flexibility was allowed for the NDNS RP participants in their choice of which day to begin their diary. But the above mentioned alternative models require that the 3 or 4 repeated measurements of food diaries to be recorded at the same time points longitudinally. Specifically, LCGA models (also called growth mixture modelling) will also need to model the change of the odds of the probability over time as a function (quadratic, or cubic) which is apparently not the objective in the current project.

Secondly, the classification of individuals to latent carbohydrate eating classes were defined by maximum posterior probability assignment rule. This approach assigns individuals to the class for which they have the highest posterior probability of membership \parencite{nagin2005group}. The other approach is multiple pseudo-class draws \parencite{wang2005residual}, which was proposed in attempt to account for uncertainty in class assignment. However, the maximum probability rule is still believed to be able to minimize the number of incorrect assignments \parencite{goodman20071}. A Monte Carlo simulation study \parencite{bray2015eliminating} also demonstrated that maximum probability assignment is less biased than multiple pseudo-class draws. Moreover, this simulation study also found that an inclusive LCA (i.e. LCA with covariates) would perform better than non-inclusive LCA, and has the potential to reduce bias of class assignment. However, whether this advantage can be extended to MLCA in the current project is unknown, and also due to the reason that the associations between the latent classes and distal outcomes (hypertension, and BMI) in our analyses are mostly still under exploration, therefore, the model for LCA with covariates can never be well-known before our analyses. Thus, as a first step, given the complexity of the NDNS RP dataset itself, we chose to conduct the MLCA modellings without any covariate in either day level or individual level models. Future studies may be advised to consider incorporating other predictors in the MLCA models to see whether the classifications in both levels can be improved or not.

Thirdly, information regarding the detailed occupations of the participants was not available and because of which we could not exclude those who had a shift work involving night work. But considering that completing a 4-consecutive-day food diary would be a large burden for the participants, those had a shift work involving night work might possibly performed relatively lower compliance and could not complete their diary for at least 3 days. But this hypothesis cannot be verified in the NDNS RP dataset. If people working with shift work were included in the NDNS RP sample, the classification in both levels as well as their associations with either hypertension or BMI would be biased. Furthermore, under-reporting due to the burden of writing down everything they eat at each occasion is also a concern in the food diaries collected. To evaluate the influence of under-reporting in the NDNS RP sample, study team of NDNS RP conducted doubly labelled water (DLW) sub-study within the survey. The details of this sub-study can be accessed from the Appendix X of the Official Reports provided by Public Health England \parencite{bates2014national,roberts2018national,NDNSofficial}. Briefly, in healthy adults, if, for a given period of time, energy consumed in food matches total energy expended, they are in energy balance. In NDNS RP sub-study, estimates of energy intake (EI) from the four-day diary were compared with measurements of total energy expenditure (TEE) using the DLW technique in a sub-sample of survey participants. Results of the sub-study showed reasonable agreement between EI and TEE (overall ratio: 0.73). Reasons such as misreporting of actual consumption; under-reported or modified usual intake due to the burden of the survey/DLW sub-study are potential factors may contributed to under-reporting of their EI. We cannot extrapolate the estimated under-reporting to the whole sample since other individuals' diet might be affected by under-reporting differentially.

Lastly, it should be noted that findings from data-driven, exploratory methods may not be
generalizable to populations in other countries. Since the carbohydrate eating patterns in both day level and individual level probably only represent the socio-cultural and life-style characteristics of the adult population in the UK. Further researches are warranted to explore and better understand the eating patterns in other populations. Besides, the cross-sectional study design also refrained us from deducing any causal effect between the carbohydrate eating patterns and hypertension or BMI.

@@

We have successfully defined carbohydrate eating patterns in the general population in the UK adults using the NDNS RP database in both observation day level and individual level. Low carbohydrate eaters turned to have more energy that contributed by both fat and alcohol. Moderate carbohydrate eaters consumed the lowest total energy, while they had the tendency of having meals later in time-of-day. High carbohydrate eaters consumed most of their carbohydrate as well as energy earlier in time-of-day. These dietary patterns specifically for carbohydrate intake were found to be differed by timing, quantity and contributions to energy consumption. Compared with low carbohydrate eaters, men had moderate carbohydrate eating pattern may associated with a lower prevalence of hypertension, women in this latent class who lived alone may associated with a larger waist circumferences. Among women who lived with partners, high carbohydrate eating pattern was associated with both lower BMI and smaller waist circumferences. Longitudinal designed studies are needed in finding whether the eating patterns themselves are changing over time and in investigating how such circadian eating patterns may relate to the change of blood pressure, obesity, and even mortality.

### 2018-08-19 paper/main.tex
appendix C figlevel 1 CW4
limitation 1
before go to chaed party
updated discussion done

\large \textit{Submitted in part fulfillment of the requirements\\ for the degree of \degreename}\\[0.3cm] % University requirement text
%\textit{in the}\\[0.4cm]
\groupname\\\deptname\\[1.8cm] % Research group name and department name

@@

\begin{abstract}
\addchaptertocentry{\abstractname} % Add the abstract to the table of contents
The National Diet and Nutrition Survey (NDNS) database of detailed four-day food diaries was used to \dots
\end{abstract}

@@

\textbf{AIC} & \textbf{A}kaike \textbf{I}information \textbf{C}riterion \\
\textbf{A1c} & Haemoglobin \textbf{A1c}: Glycated hemoglobin \\
%\textbf{aBIC} & \textbf{a}djusted \textbf{B}ayesian \textbf{I}nformation \textbf{C}riterion \\
\textbf{BMI} & \textbf{B}ody \textbf{M}ass \textbf{I}ndex \\
%\textbf{cAIC} & \textbf{c}onsistent \textbf{A}kaike \textbf{I}nformation \textbf{C}riterion \\
\textbf{BIC} & \textbf{B}ayesian \textbf{I}information \textbf{C}riterion \\
\textbf{CI}  & \textbf{C}onfidence \textbf{I}nterval \\
\textbf{DM}  & \textbf{D}iabetes \textbf{M}ellitus \\
%\textbf{EM}  & \textbf{E}xpectation \textbf{M}aximazation  \\
\textbf{FSA} & \textbf{F}ood \textbf{S}tandards \textbf{A}gency  \\
\textbf{HDL} & \textbf{H}igh \textbf{D}ensity \textbf{L}ipoprotein cholesterol \\
\textbf{LDL} & \textbf{L}ow \textbf{D}ensity \textbf{L}ipoprotein cholesterol \\
\textbf{LCA} & \textbf{L}atent \textbf{C}lass \textbf{A}nalysis\\
%\textbf{LCGA} & \textbf{L}atent \textbf{C}lass \textbf{G}rowth \textbf{A}nalysis \\
%\textbf{LTA} & \textbf{L}atent \textbf{T}ransition \textbf{A}nalysis \\
\textbf{MAFF} & \textbf{M}inistry of \textbf{A}griculture, \textbf{F}isheries and \textbf{F}ood \\
%\textbf{MAR} & \textbf{M}issing \textbf{A}t \textbf{R}andom \\
%\textbf{MCAR} & \textbf{M}issing \textbf{C}ompletely \textbf{A}t \textbf{R}andom \\
\textbf{MLCA} & \textbf{M}ultilevel \textbf{L}atent \textbf{C}lass \textbf{A}nalysis \\
%\textbf{MNAR} & \textbf{M}issing \textbf{N}ot \textbf{A}t \textbf{R}andom \\
\textbf{ML} & \textbf{M}aximum \textbf{L}ikelihood \\
\textbf{NDNS} & the \textbf{N}ational \textbf{D}iet and \textbf{N}utrition \textbf{S}urvey \\
\textbf{OR} & \textbf{O}dds \textbf{R}atio \\
\textbf{PHE} & \textbf{P}ublic \textbf{H}ealth \textbf{E}ngland  \\
\textbf{PSUs} & \textbf{P}rimary \textbf{S}ampling \textbf{U}nits \\
\textbf{TC} & \textbf{T}otal \textbf{C}holesterol \\
\textbf{TG} & \textbf{T}ri\textbf{g}lycerides \\
\textbf{WC}   & \textbf{W}aist \textbf{C}ircumference \\
\end{abbreviations}

@@

\include{Appendices/AppendixA}
\include{Appendices/AppendixB}
\include{Appendices/AppendixC}

>>>>

\large \textbf{Submitted in part fulfillment of the requirements\\ for the degree of \degreename}\\[0.3cm] % University requirement text
%\textit{in the}\\[0.4cm]
\groupname\\\deptname\\[1.8cm] % Research group name and department name

@@

\begin{abstract}
\addchaptertocentry{\abstractname} % Add the abstract to the table of contents
The National Diet and Nutrition Survey (NDNS) database of detailed four-day food diaries was used in exploration of both temporal and quantity eating patterns of carbohydrates in the general adults living in the UK. 6155 participants aged 19 or older were kept in the analysis. \\ 
Multilevel latent class analysis (MLCA) models were applied in finding the latent classes of carbohydrate consumption both at the observation day level, and individual level.
\end{abstract}

@@

\textbf{AIC} & \textbf{A}kaike \textbf{I}information \textbf{C}riterion \\
\textbf{A1c} & Haemoglobin \textbf{A1c}: Glycated hemoglobin \\
%\textbf{aBIC} & \textbf{a}djusted \textbf{B}ayesian \textbf{I}nformation \textbf{C}riterion \\
\textbf{BMI} & \textbf{B}ody \textbf{M}ass \textbf{I}ndex \\
%\textbf{cAIC} & \textbf{c}onsistent \textbf{A}kaike \textbf{I}nformation \textbf{C}riterion \\
\textbf{BIC} & \textbf{B}ayesian \textbf{I}information \textbf{C}riterion \\
\textbf{CI}  & \textbf{C}onfidence \textbf{I}nterval \\
\textbf{DM}  & \textbf{D}iabetes \textbf{M}ellitus \\
%\textbf{EM}  & \textbf{E}xpectation \textbf{M}aximazation  \\
\textbf{FSA} & \textbf{F}ood \textbf{S}tandards \textbf{A}gency  \\
\textbf{HDL} & \textbf{H}igh \textbf{D}ensity \textbf{L}ipoprotein cholesterol \\
\textbf{LDL} & \textbf{L}ow \textbf{D}ensity \textbf{L}ipoprotein cholesterol \\
\textbf{LCA} & \textbf{L}atent \textbf{C}lass \textbf{A}nalysis\\
\textbf{LCGA} & \textbf{L}atent \textbf{C}lass \textbf{G}rowth \textbf{A}nalysis \\
\textbf{LTA} & \textbf{L}atent \textbf{T}ransition \textbf{A}nalysis \\
\textbf{MAFF} & \textbf{M}inistry of \textbf{A}griculture, \textbf{F}isheries and \textbf{F}ood \\
%\textbf{MAR} & \textbf{M}issing \textbf{A}t \textbf{R}andom \\
%\textbf{MCAR} & \textbf{M}issing \textbf{C}ompletely \textbf{A}t \textbf{R}andom \\
\textbf{MLCA} & \textbf{M}ultilevel \textbf{L}atent \textbf{C}lass \textbf{A}nalysis \\
%\textbf{MNAR} & \textbf{M}issing \textbf{N}ot \textbf{A}t \textbf{R}andom \\
\textbf{ML} & \textbf{M}aximum \textbf{L}ikelihood \\
\textbf{NDNS} & the \textbf{N}ational \textbf{D}iet and \textbf{N}utrition \textbf{S}urvey \\
\textbf{OR} & \textbf{O}dds \textbf{R}atio \\
\textbf{PHE} & \textbf{P}ublic \textbf{H}ealth \textbf{E}ngland  \\
\textbf{PSUs} & \textbf{P}rimary \textbf{S}ampling \textbf{U}nits \\
\textbf{RMLCA} & \textbf{R}epeated \textbf{Measures} \textbf{LCA} \\
\textbf{TC} & \textbf{T}otal \textbf{C}holesterol \\
\textbf{TG} & \textbf{T}ri\textbf{g}lycerides \\
\textbf{WC}   & \textbf{W}aist \textbf{C}ircumference \\
\end{abbreviations}

@@

\include{Appendices/AppendixA}
\include{Appendices/AppendixB}
\include{Appendices/AppendixC}
\include{Appendices/AppendixD}
\include{Appendices/AppendixE}
